# Git

```{r pkg1, include = FALSE, R.options = list(tidyverse.quiet = TRUE)}
sapply(
  c(
    'kableExtra',
    'magrittr'
  ),
  require,
  character.only = TRUE
) 
```

[Git](https://git-scm.com/) is a free and open-source version control system. When used regularly and as intended, developers will have a full history of their project within a local repository. In addition to a historical log of changes to project files, git allows for project branching to support users to test/develop new code, while maintaining the master version for easy reversion. When users are ready to implement their new branch into the main codebase git can be used for merging files, whereby it tracks changes and ensures there are no conflicts between the main version and the new version. 

Beyond local usage, git is also supported by web-based repositories such as [Github](https://github.com/) and [Gitlab](https://about.gitlab.com/), where projects can be pushed, pulled, and cloned. These sites allow for easy collaboration with other developers and provide a number of user-friendly features that make working with git easier.

:::: {.notebox}

::: {.center}
**NOTE**
:::

In RStudio, git is accessed in the terminal or in the top right under the Git tab when Git has been initialized for the project.

::::

## Basic git workflow {#basic-workflow}

Step 1) Create a project.

Step 2) Initialize git.   
`git init`

Step 3) Check your project file status
`git status`

Step 4) Add a file to the local repo   
`git add filename.R`

Step 5) Commit your change with a message   
`git commit -m "Add filename.R"`

Step 6) Create a branch to test/develop code   
`git branch test_branch`

Step 7) Go into that branch   
`git checkout test_branch`

Step 8) Modify your code and repeat steps 3 and 4

Step 9) When ready merge branches     
`git merge master test_branch`

:::: {.tipbox}

::: {.center}
**TIP**
:::

Branching is useful when you have a stable codebase that you do not want to break or if you are working collaboratively on a code base. You can create a branch to do development work and test new features until it's ready for integration with the stable codebase or your coworkers. If you are developing something from scratch by yourself you may not need to use branching until later.

::::

## Setting up Github or Gitlab projects

Ensure your profile is set up   
`git config --global user.name "Braeden Klaver"`   
`git config --global user.email "braeden.klaver@bccdc.ca"`   

::::{.warnbox}
:::{.center}
**WARNING**
:::
When you create a new repository on github or gitlab it can cause issues if you create it with a README. It is recommended when creating new repositories to make sure they are a clean slate.
::::

### Pre-existing project on Github or Gitlab

Step 1) In your terminal navigate to the folder you want to clone the project to   
`cd "U:/myprojects"`

Step 2) Clone the project into that folder and give the project a name   
`git clone http://lvmgenodh01.phsabc.ehcnet.ca/braeden.klaver/test.git myproject` 

### Personal project without git

Step 1) Open your R project

Step 2) Initialize git   
`git init`

Step 3) Create a project on [Gitlab](https://about.gitlab.com/) or [Github](https://github.com/)

Step 4) Connect your R project to that repository (should be the URL)   
`git remote add origin http://lvmgenodh01.phsabc.ehcnet.ca/braeden.klaver/test.git`

Step 5) Add and commit your files   
`git add .`   
`git commit -m "Initial commit"`

Step 6) Push your project to that repository    
`git push -u origin --all`


### Personal project with Git

Step 1) Open your R project

Step 2) Connect your R project to that repository (should be the URL)   
`git remote add origin http://lvmgenodh01.phsabc.ehcnet.ca/braeden.klaver/test.git`

Step 3) Push your project to that repository   
`git push -u origin --all`

## Collaborative git workflow

In addition to the basic workflow, when working with a web-based repository like [Github](https://github.com/) or [Gitlab](https://about.gitlab.com/) there are additional steps you will need to take:

Step 1) Pulling changes from the repository - your coworkers may have made changes!   
`git pull`

Step 2) Follow the basic workflow (Section \@ref(basic-workflow))

Step 3) Pushing your changes to the repository - your coworkers will want to be up to date!   
`git push` or `git push origin` 

::::{.notebox}
:::{.center}
**NOTE**
:::
You have to use `git push origin branch_name` when working on a branch that isn't your main or master.
::::

## Using Git at the BCCDC

### Git on Citrix

[Git](https://git-scm.com/) is installed on citrix already, it can be initialized as described above or you can click a check box to create a git repository when creating a new project.

:::: {.tipbox}

::: {.center}
**TIP**
:::

If you have a local R installation and you'd like to work there but do not have git installed locally you can still leverage citrix R to use git for your projects.

::::

### Git locally

If you want to use [Git](https://git-scm.com/) locally you will need to install it for your system. Typically, this will be the 64-bit windows version found [here](https://git-scm.com/downloads/win). Once it is downloaded you should be able to use git as described above in your local RStudio.

:::: {.notebox}
:::{.center}
**NOTE**
:::
The benefits of having a local installation of git is that if you primarily use your local RStudio you can access it directly through there rather than jumping to citrix. Additionally, you can use it to connect to repositories in Github, which is not possible on citrix.
::::

### Github

Because [Github](https://github.com/) is an external web-based repository there are some considerations for its use at the BCCDC. Currently, there are no formal guidelines on using it, for this reason using it for BCCDC-specific projects should be avoided.


### Gitlab

The BCCDC has a private [Gitlab repository](http://lvmgenodh01.phsabc.ehcnet.ca/), which can be used for regular BCCDC projects within the scope set out in the [guiding document](https://healthbc-my.sharepoint.com/:w:/r/personal/kathleen_mclean_bccdc_ca/Documents/GitLab%20Guideline%20for%20distribution.docx?d=w2584389db5be4a7d8dc59ed28080e351&csf=1&web=1&e=OH30gE). 

:::: {.notebox}

::: {.center}
**NOTE**
:::

You can request access to Gitlab via this [form](https://surveys.vch.ca/Survey.aspx?s=62de2ddc1daf413091a8e063d80428cd).

::::

### Suggested project workflow

Working with git tracked projects requires you to have your own local repository. For this reason, it is recommended to keep this repository in your `U:/` drive. In addition, we would recommend having another local repository in the `O:/` drive, which would be dedicated to running pipelines or deploying apps and **not** for development work.

:::: {.warnbox}

::: {.center}
**WARNING**
:::

Some project data may not be permitted on your `U:/` drive, therefore ensure your code is loading that data from an appropriate location.

::::

## Basic Git Commands Summary

``` {r gitcommands, echo=F}
data.frame(
  Command = c(
    'git init',
    'git status',
    'git add',
    'git rm',
    'git commit',
    'git branch',
    'git checkout',
    'git merge',
    'git pull',
    'git push'
  ),
  Description = c(
    'Initialize git for the directory',
    'Check the status of files in the directory (eg. are they being tracked, have they been modified)',
    'Stage a file for commit',
    'Remove a file staged for commit',
    'Commit your changes',
    'Create a branch for development work',
    'Checkout a branch to work within',
    'Merge two branches together',
    'Pull the most up-to-date repository from a remote (ie. Github or Gitlab)',
    'Push your changes to a remote repository (ie. Github or Gitlab)'
  )
) %>% 
  knitr::kable()
```

:::: {.notebox}

::: {.center}
**NOTE**
:::

There are many other functions that you can use in git beyond those listed above, however, these give you the tools to get started.

::::

## Additional Resources

- [Mike Kuo's Git Tutorial](https://healthbc-my.sharepoint.com/:u:/g/personal/michael_kuo_bccdc_ca/EbDhRI_pKvJAlISoY3xaEGUBIwMN2bwzvraOdvK4OZTriQ)   
- [Git Playground](https://learngitbranching.js.org/)   

:::: {.notebox}
:::{.center}
**NOTE**
:::
The Git Playground does not work on Microsoft Edge! Try it on Google Chrome.
::::

