# Shiny

```{r, include = FALSE, R.options = list(tidyverse.quiet = TRUE)}
sapply(
  c(
    'shiny'
  ),
  require,
  character.only = TRUE
) 
```

If you’ve never used Shiny before, welcome! Shiny is an R package that allows you to easily create rich, interactive web apps. Shiny allows you to take your work in R and expose it via a web browser so that anyone can use it. Shiny makes you look awesome by making it easy to produce polished web apps with a minimum amount of pain.

## Shiny App Structure

Shiny apps are composed of two main elements.

- The UI: This is where you define the layout and appearance of your app, including sliders, plots, tabs, etc.
- The server: This is where you connect your UI components using logic behind the scenes to drive app behaviour.

These two elements are operationalized by calling `shinyApp(ui = ui, server = server)`

## UI

### Setting up the UI

TO create a shiny app you must create a UI. Traditionally, using shiny you would do this using the function `fluidPage()`. The UI will contain elements such as calls to CSS stylings, overall UI design, inputs, and outputs.

### Layout and themes



al lof the page layouts and panels, sidebars

CSS

html

### Inputs

There are a number of inputs that are incredibly useful in shiny apps such as `radioButtons()`, `selectInput()`, `actionButton()`, and `dateRangeInput()`. These allow users to interact with our app to dictate what appears in our app.

These functions all share a two main arguments:

- `inputID`: This is the identifier used to connect the front end with the back end: if your UI has an input with ID "name", the server function will access it with input$name.   
- `label`: This is used to create a human-readable label for the control (ie. Select Geography). 

:::: {.notebox}

::: {.center}
**NOTE**
:::

The `inputId` has two constraints:

- It must be a simple string that contains only letters, numbers, and underscores (no spaces, dashes, periods, or other special characters allowed!). Name it like you would name a variable in R.

- It must be unique. If it’s not unique, you’ll have no way to refer to this control in your server function!

::::

### Outputs

In the UI we can also specify the outputs that we'd like to include, for example plots, text, or tables. In essence, these are 'placeholders', which will be filled in based on what we define in our server. Similar to inputs, we must also specify an `inputID` as the first argument.


## Server

On the server side, we need to use a different suite of functions that will focus on three main things:

- Inputs: these are referenced by calling their IDs using `input$button1` for example.
- Outputs: Because we set up our 'placeholders' in the UI we need to tell them what to be filled with here (ie. render a plot).
- Reactivity: these will determine the behaviour of your app, typically in response to things like `input$button1`.

## Dynamic UI

## Basic example



## Deploying your app

## Packaging a Shiny App

Using package structure for a shiny app gets your toes into the water of package development. It’s a long way from a complete package, but it’s still useful because it activates new tools that make it easier to work with larger app and provides you with standard conventions that can be used across projects. 

### Put all R code in the `R/` directory

Because we are going to be working in an app-package we need to create an `R/` directory. This is where we will keep all of the core R code components that will build our app. As a reminder, in a package we are leveraging useful tools like `devtools::load_all()` (See section \@ref(load)), which will go into the `R/` directory and load/run all of the code within this directory.

### Write a function that starts your app

As mentioned in section (LINK HERE), we need three primary pieces to create an app:

``` {r eval=F}
ui <- fluidPage(
  ...
)

server <- function(input, output, session) {
  ...
}

shinyApp(ui, server, ...)
```

Now, to begin converting our project to a package-app we must wrap this into a function:

``` {r eval=F}
myApp <- function(...) {
  ui <- fluidPage(
    ...
  )
  server <- function(input, output, session) {
    ...
  }
  shinyApp(ui, server, ...)
}
```

:::: {.notebox}

:::{.center}
**NOTE**
:::

The function you create shouldn't need any arguments, all that we will use this for is to easily run our app locally and set it up for easy deployment later.

::::

### Save your data to the `data/` directory

We may have some datasets or lists that we use in our app that aren't routinely updated. These are perfect candidates to convert into `.rda` files using the handy function `usethis::use_data()` that we talked about in section \@ref(pkg-data). These will automatically be stored in the `data/` directory and can be easily called directly after running `load_all()`.

### Create an `inst/` directory

The `inst/` directory is where we can store other raw datasets that are more subject to change, for example the data used to feed our apps that comes from external pipelines. There is no standard convention so you can name things as you please, but common usage includes folders named `inst/extdata` or `inst/ext` to store these datasets.

:::: {.notebox}

:::{.center}
**NOTE**
:::

Load data with `read.csv(system.file("exdata", "mydata.csv", package = "myApp"))`. Notice that our package-app automatically knows to look in our `inst/` folder?

``` {r eval=F}
myApp <- function(...) {
  read.csv(system.file("extdata", "mydata.csv", package = "myApp"))
  ui <- fluidPage(
    ...
  )
  server <- function(input, output, session) {
    ...
  }
  shinyApp(ui, server, ...)
}
```

::::

### Create a `www/` directory

This directory is where you can store some of the other non-script or data components to the app such as CSS stylings or images. There are no rules to this but a suggestion would be to have a `www/css` and `www/images` sub-directory scheme for your app.

:::: {.tipbox}

:::{.center}
**TIP**
:::

If you want to call files from within these sub-directories (eg. `href = "css/style.css"`) you will need to tell your app where they are:

```{r eval=F}
myApp <- function(...) {
  shiny::addResourcePath("css", file.path(getwd(), "www/css"))
  shiny::addResourcePath("images", file.path(getwd(), "www/images"))
  ui <- fluidPage(
    ...
  )
  server <- function(input, output, session) {
    ...
  }
  shinyApp(ui, server, ...)
}
```

::::

### Create a `DESCRIPTION` file

Because we are working in a package environment one of the critical components will be our `DESCRIPTION` file, which was discussed in section \@ref(desc). Don't forget to set your dependencies!


### Deploying your app-package

One of the final pieces to setting up your app-package is the `app.R` script. This is what will be used when we go to deploy our app to the server and contains two simple but important lines of code:

```{r eval=F}
pkgload::load_all(".")
myApp()
```

:::: {.warningbox}

:::{.center}
**WARNING**
:::

Although this is an R script, we **DO NOT** place this under our `R/` directory. This would result in an infinite loop when loading due to our `load_all()` function! Therefore place this at the top-level of your project directory.

::::

Normally when you deploy an app, the rsconnect package automatically figures out all of the packages your code uses. But now that you have a `DESCRIPTION` file, it requires you to explicitly specify them. The easiest way to do this is to call usethis::use_package(). You’ll need to start with shiny and pkgload:   

```{r eval=F}
usethis::use_package("shiny")
usethis::use_package("pkgload")
```

[Read more!](https://mastering-shiny.org/scaling-packaging.html#deploying-your-app-package)

### Workflow

Putting your app code into the package structure unlocks a new workflow:

- Re-load all code in the app with `Cmd/Ctrl + Shift + L`. This calls `devtools::load_all()` which automatically saves all open files, `source()`s every file in `R/`, loads all datasets in `data/` then puts your cursor in the console.  

- Re-run the app with `myApp()`.   

[Read more!](https://mastering-shiny.org/scaling-packaging.html#workflow)

## Additional Resources

- [Shiny Manual](https://mastering-shiny.org/)   
- [Shiny Cheatsheet](https://rstudio.github.io/cheatsheets/html/shiny.html)

