# Modularization

At the simplest level, a module is a pair of UI and server functions. The magic of modules comes because these functions are constructed in a special way that creates a “namespace”. So far, when writing an app, the names (ids) of the controls are global: all parts of your server function can see all parts of your UI. Modules give you the ability to create controls that can only be seen from within the module. This is called a namespace because it creates “spaces” of “names” that are isolated from the rest of the app.

Shiny modules have two big advantages:   

- Namespacing makes it easier to understand how your app works because you can write, analyse, and test individual components in isolation.     
- Modules are functions that help you reuse code; anything you can do with a function, you can do with a module.    

## Module Basics

Unlike regular Shiny code, connecting modules together requires you to be explicit about inputs and outputs. Initially, this is going to feel tiresome and it’s certainly more work than Shiny’s usual free-form association. But modules enforce specific lines of communication for a reason: they’re a little more work to create, but much easier to understand, and allow you to build substantially more complex apps.

To showcase modularization let's look at a very simple app from [mastering shiny](https://mastering-shiny.org/scaling-modules.html#module-basics):

``` {r eval=F}

ui <- shiny::fluidPage(
  shiny::selectInput("var", "Variable", names(datasets::mtcars)),
  shiny::numericInput("bins", "bins", 10, min = 1),
  shiny::plotOutput("hist")
)

server <- function(input, output, session) {
  data <- shiny::reactive(datasets::mtcars[[input$var]])
  output$hist <- shiny::renderPlot({
    hist(
      data(), 
      breaks = input$bins, 
      main = input$var
    )
  }, 
  res = 96
  )
}

```

As we begin to modularize this app you'll notice that a module is very similar to an app, that is, they are composed of two parts:    

- A module UI function
- A module server function

These functions both require an `id` argument that will namespace the module when we use it in our app. To begin creating a module you are going to look for UI/server pairs in your app and extract them out into paired UI/server functions! In our example, this will be our inputs and plot.

[Read more!](https://mastering-shiny.org/scaling-modules.html#module-basics)

::::{.notebox}
:::{.center}
**NOTE**
:::
You can test your modules by treating them like a simple standalone app!
::::

### Module UI

First let's set up the UI module. There are two steps:   

- Put the UI in a function with an `id` argument   
- Wrap each existing `id` in a `NS()` function so that each previous `id` turns into `NS(id, "previous_id")`

This would look like this for our example:

``` {r eval=F}
histogramUI <- function(id) {
  htmltools::tagList(
    shiny::selectInput(shiny::NS(id, "var"), "Variable", choices = names(mtcars)),
    shiny::numericInput(shiny::NS(id, "bins"), "bins", value = 10, min = 1),
    shiny::plotOutput(shiny::NS(id, "hist"))
  )
}
```

The UI components are wrapped in `tagList()`, which allows you to bundle components that can then be placed in other functions in the app code such as `shiny::fluidRow()`, `bslib::layout_column_wrap()` or `bslib::sidebar()`. You can also wrap them in `shiny::fluidRow()` or `bslib::layout_column_wrap()` in the function itself if you'd prefer, though that can limit the reusability.

[Read more!](https://mastering-shiny.org/scaling-modules.html#module-ui)

::::{.notebox}
:::{.center}
**NOTE**
:::
There is another way to namespace within a UI function though this does add extra steps and can make calling these namespaces in the server a little more tricky:

``` {r eval=F}

histogramUI <- function(id) {
  ns <- shiny::NS(id)
  shiny::plotOutput(
    ns("hist")
  )
}

```
::::

### Module server

The second part of the module is the server function, which is also going to have an `id` argument; we will use the `id` to link it to our UI component later when building our app. Inside of this function we call `moduleServer()` with the `id` arugment, and an additional function that looks like a typical server function:

``` {r eval=F}
# your first line will look like this, you may have more
# arguments, we'll look at that later
histogramServer <- function(id) { 
  # your second line will always look like this
  shiny::moduleServer(id, function(input, output, session) {
    # you put your server code inside of this call
    data <- shiny::reactive(datasets::mtcars[[input$var]])
    output$hist <- shiny::renderPlot({
      hist(data(), breaks = input$bins, main = input$var)
    }, res = 96)
  })
}
```

These two levels help distinguish the arguments between your module and the server function. They may look complex but this is how every module is set up, so it's really a copy and paste scenario.

[Read more!](https://mastering-shiny.org/scaling-modules.html#module-server)

::::{.notebox}
:::{.center}
**NOTE**
:::
`shiny::moduleServer()` takes care of the namespacing automatically, that is, we can refer to `NS(id, "var")` and `NS(id, "bins")` from the UI section of the module by calling `input$var` and `input$bins`. This is because of the shared `id` argument between the functions, that we will connect when we implement them in our app like usual.
::::

### Module testing

It's good practice to test your module in a function that creates a simple app:

``` {r eval=F}
histogramApp <- function() {
  ui <- shiny::fluidPage(
    histogramUI("hist1")
  )
  server <- function(input, output, session) {
    histogramServer("hist1")
  }
  shiny::shinyApp(ui, server)  
}
```

::::{.notebox}
:::{.center}
**NOTE**
:::
Just like a regular app we need to link the UI and server using the same `id` name. This will be what allows the UI and server parts of the module to talk to eachother and identify the inputs and outputs. 
::::

### Namespacing

Let's circle back on namespacing. At this point of the example we have two separate namespacing situations:

- Inside of our module we have `NS(id, "var")`, `NS(id, "bins")`, and `NS(id, "hist")`
- Outside of our module in our app function we have `histogramUI("hist1")` and `histogramServer("hist1")`

It's important to realize here that the namespaces that are created inside a module only exist (ie. can be referred to) within the two module functions; in other words, a module is in essence a "black box" that can't be seen from outside of the module. This means that as an author of a module, you don't need to worry about conflicts with namespacing with other modules or with namespacing in the app UI or server.

Even though modules are a "black box" you can still use outside inputs by adding additional function arguments or module outputs to drive functionality with other modules or server elements - we'll talk about this later in section \@ref(advanced-modules).

[Read more!](https://mastering-shiny.org/scaling-modules.html#namespacing)

::::{.notebox}
:::{.center}
**NOTE**
:::
Note that the module UI and server differ in how the namespacing is expressed:   

- In the module UI, the namespacing is explicit: you have to call `NS(id, "name")` every time you create an input or output.   
- In the module server, the namespacing is implicit. You only need to use `id` in the call to `moduleServer()` and then Shiny automatically namespaces `input` and `output` so that in your module code `input$name` refers to the input with `NS(id, "name")`.   
::::

### Single object modules

When some people encounter modules for the first time, they may attempt to combine the module server and module UI into a single module object. However, in Shiny, UI and server are inherently disconnected; Shiny doesn’t know which UI invocation belongs to which server session. You can see this pattern throughout Shiny: for example, `plotOutput()` and `renderPlot()` are connected only by shared `id`. Writing modules as separate functions reflects that reality: they’re distinct functions that are not connected other than through a shared `id`. This also allows you to make them more generalizable and allows for reactivity.

For a full example [read more here](https://mastering-shiny.org/scaling-modules.html#module-objects).

### Module naming conventions

A standard convention for naming can be as follows:

- `R/histogram.R` holds all the code for the module.    
- `histogramUI()` is the module UI. If it’s used primarily for `input` or `output` it can be called `histogramInput()` or `histogramOuput()` instead.    
- `histogramServer()` is the module server.   
- `histogramApp()` creates a complete app for interactive experimentation and more formal testing.   

[Read more!](https://mastering-shiny.org/scaling-modules.html#naming-conventions)

## Advanced Modules {#advanced-modules}

Adding arguments beyond the `id` to the module UI and server gives greater control over the module, allowing you to use the same module in more places in your app. Further, you can return one or more reactive values from your server module so that you can use them in your larger app server call.

[Read more!](https://mastering-shiny.org/scaling-modules.html#inputs-and-outputs)

### Additional UI arguments

Let's look at an example from the [book](https://mastering-shiny.org/scaling-modules.html#getting-started-ui-input-server-output) where we will have an additional argument in our UI function. We do it the same way as any other function, we add it as an argument in our function call:

``` {r eval=F}
# Two arguments, id as usual and a filter
datasetInput <- function(id, filter = NULL) {
  
  # Get list of df names for user selection
  names <- ls("package:datasets")
  
  # If a filter is present then update the names in the list
  # accordingly
  if (!is.null(filter)) {
    # get the data to identify datasets vs matrices
    data <- lapply(names, get, "package:datasets")
    # filter for selection, "dataset" or "matrix"
    names <- names[vapply(data, filter, logical(1))]
  }
  
  # Set up the UI selections for the list of names
  shiny::selectInput(shiny::NS(id, "dataset"), "Pick a dataset", choices = names)
  
}
```

### Server outputs

Now let's look at the server function. When we are planning to use an output from a server module we will need to wrap it in a `shiny::reactive()` as it's technically a reactive output given it's relationship with the UI. The example sets up the server like so:

``` {r eval=F}
datasetServer <- function(id) {
  # set up the usual 
  shiny::moduleServer(id, function(input, output, session) {
    # The last expression in the function
    # will be the return value and return values
    # should always be inside a reactive() wrapper
    shiny::reactive(get(input$dataset, "package:datasets"))
  })
}
```

### Operationalizing advanced modules

Here we put the two parts together. The UI portion is fairly simple, with the only difference being the `filter` argument was included in the larger app call as well. However, the server part is a bit more complicated. When you are returning a value in the server portion of your app from your module you must assign it to a value in the environment, just like you would in typical R code (ie. `x <- mymodule_server()`). Because we are working with shiny, and the returned values from modules are reactive, the challenge is deciding if you want to refer to the reactive aspect of the value or the value itself.. let me explain:

- When you want to use that stored value/object (ie. the dataset or value itself) you must refer to it as `value()` or `data()` for example.
- When you want to use the reactive aspect to drive more server logic/behaviour you would just refer to `value` or `data`

Let's see how this looks in the example:

``` {r eval=F}

# They included an argument in the larger app call
# it makes sense for this small example but you would probably
# call it in the UI itself.. at least I would
datasetApp <- function(filter = NULL) {
  ui <- shiny::fluidPage(
    # Set up the input portion of the module
    datasetInput("dataset", filter = filter),
    # also include a placeholder for a table to show
    # the selected data from our module
    shiny::tableOutput("data")
  )
  server <- function(input, output, session) {
    # Set up the server, because it's giving us a reactive value
    # that we want to use in the table we need to do `<-` assignment
    # just like we would normally in R
    data <- datasetServer("dataset")
    # Now call this data in our table render
    output$data <- renderTable(head(data()))
  }
  shinyApp(ui, server)
}
```

[Read more!](https://mastering-shiny.org/scaling-modules.html#getting-started-ui-input-server-output)

::::{.notebox}
:::{.center}
**NOTE**
:::
In the above example it called the module output as `data()` when rendering it into the table because we wanted to use the dataset to create our table and **not** the reactive aspect of the data. The reactive forms are often used in other module calls or in wrappers like `shiny::observeEvent()`.
::::

### Server inputs

Sometimes we want to use other module outputs or outputs created in our larger server call in a module. We can do this by adding an argument in our top level server function. We also need to consider what kind of value we want to use for that argument (ie. reactive versus the actual values); for this reason it can be a good idea to set up simple error messages in your server modules to ensure that future users know what type of value is expected in a server argument:

``` {r eval=F}
selectVarServer <- function(id, data, filter = is.numeric) {
  # We need data to be reactive!
  stopifnot(shiny::is.reactive(data))
  # We need the filter to be a non-reactive value!
  stopifnot(!shiny::is.reactive(filter))
  
  shiny::moduleServer(id, function(input, output, session) {
    shiny::observeEvent(data(), {
      # See that we call data() here and just filter
      shiny::updateSelectInput(session, "var", choices = find_vars(data(), filter))
    })
    
    shiny::reactive(data()[[input$var]])
  })
}
```

[Read more!](https://mastering-shiny.org/scaling-modules.html#server-inputs)

::::{.notebox}
:::{.center}
**NOTE**
:::
Another way to think about reactive or non-reactive values is when can the value change: is it fixed and constant over the life-time of the app, or is it reactive, changing as the user interacts with the app.
::::

### Multiple server outputs

Sometimes it's useful for the module's server to return more than one value. You can easily do this the same way you would for a regular R function - return a list. 
The example from [mastering shiny](https://mastering-shiny.org/scaling-modules.html#multiple-outputs) is:

``` {r eval=F}

selectVarServer <- function(id, data, filter = is.numeric) {
  stopifnot(shiny::is.reactive(data))
  stopifnot(!shiny::is.reactive(filter))
  
  moduleServer(id, function(input, output, session) {
    shiny::observeEvent(data(), {
      shiny::updateSelectInput(session, "var", choices = find_vars(data(), filter))
    })
    
    list(
      # Return the name and the values of the variable
      name = shiny::reactive(input$var),
      value = shiny::reactive(data()[[input$var]])
    )
  })
}

histogramApp <- function() {
  ui <- shiny::fluidPage(...)

  server <- function(input, output, session) {
    data <- datasetServer("data")
    # assign the list to X
    x <- selectVarServer("var", data)
    # create a histogram using the values and title it with the
    # variable name. We refer to the reactive forms of the values
    # because within this server module the values are called from the reactive
    histogramServer("hist", x$value, x$name)
  }
  shinyApp(ui, server)
} 

```

[Read more!](https://mastering-shiny.org/scaling-modules.html#multiple-outputs)

::::{.warnbox}
:::{.center}
**WARNING**
:::
The main challenge with this sort of code is remembering when you use the reactive form (e.g. `x$value`) vs. when you use its stored value (e.g. `x$value()`). Just remember that when passing an argument to a module, you want the module to react to the value changing which means that you have to pass the reactive, not it’s current value. Within the module the value can then easily be called when needed.
::::

### Modules inside of modules

Modules can also be nested, in that you can call a module within another module. This makes sense if you have two components that are inherently tied together in your app. Here is an example from [mastering shiny](https://mastering-shiny.org/scaling-modules.html#modules-inside-of-modules):

``` {r eval=F}

# Because the example's goal is to allow users to select
# a dataset and then select a variable from that dataset
# it made sense to just combine the individual modules into a larger
# module
selectDataVarUI <- function(id) {
  # Tag list for combining multiple components
  htmltools::tagList(
    # This is a dataset input selection
    datasetInput(shiny::NS(id, "data"), filter = is.data.frame),
    # This is to select the variable from the selected dataset
    selectVarInput(shiny::NS(id, "var"))
  )
}

selectDataVarServer <- function(id, filter = is.numeric) {
  shiny::moduleServer(id, function(input, output, session) {
    # Here we pull the data based on the selection
    data <- datasetServer("data")
    # and here we filter the data for the variable
    var <- selectVarServer("var", data, filter = filter)
    var #return value
  })
}

selectDataVarApp <- function(filter = is.numeric) {
  ui <- shiny::fluidPage(
    shiny::sidebarLayout(
      # Now there only needs to be a single call in the UI
      shiny::sidebarPanel(selectDataVarUI("var")),
      shiny::mainPanel(shiny::verbatimTextOutput("out"))
    )
  )
  server <- function(input, output, session) {
    # and a single call in the server 
    var <- selectDataVarServer("var", filter)
    output$out <- shiny::renderPrint(var(), width = 40)
  }
  shinyApp(ui, server)
}
```

[Read more!](https://mastering-shiny.org/scaling-modules.html#modules-inside-of-modules)

::::{.tipbox}
:::{.center}
**TIP**
:::
Nesting can be a really good idea because functionally it requires us to break down our code to the smallest possible building blocks. The two main benefits are:   

- We can easily build new larger functions with these smaller components.    
- We can test these smaller components before combining them - troubleshooting code in shiny is a bit more challenging that typical R, so this makes it easier!   
::::

## Examples from the PHMRSapp

To ground some of our knowledge let's look at examples from our very own PHMRSapp. These aren't necessarily perfect so try to think of ways you could improve them and test out your ideas! 

### Dropdown module

First let's break down a simple module that helps create a dropdown menu. We will go over the module that currently exists in the PHMRSapp and then consider ways that we could improve it according to the mastering shiny documentation.

Here is the original module:

``` {r eval=F}

# For this module we first set up the UI
# As always we include that id argument, but
# we also have a label argument so we can customize
# our menu title depending on what we're using it for
dropdown_ui <- function(id, label) {
  # Here's the alternative method, not needed for the other way
  ns <- shiny::NS(id)
  # The dropdown function from shiny
  shiny::selectizeInput(
    ns(id), # we also could've set this up by just doing shiny::NS(id, "selections")
    shiny::h5(label), # here we use the label argument to title our menu
    choices = NULL, # chioces are set to null because they will be filled later for what we need
    options = list(allowEmptyOption = FALSE) # makes it so we don't have a blank option
  )
}

# Now we set up the server end
# again we have our id argument, but we also have our 
# choice_list which we can use to update our menu options
dropdown_server <- function(id, choice_list) {
  # the usual call to moduleServer()
  shiny::moduleServer(
    id,
    function (input, output, session) {
      # We will update our choices here and set
      # the selected to be the top choice in our
      # list by default
      shiny::updateSelectizeInput(
        session,
        id,
        choices = choice_list,
        selected = choice_list[[1]][1]
      )
      # creating an empty list of reactive values that can be filled
      myreturn <- shiny::reactiveValues()
      # set up the return to txt in a reactive container
      shiny::observe({ 
        myreturn$txt <- shiny::req(eval(parse(text = paste0("input$", id))))
        })
      return(myreturn)
    })
}

# Finally, let's test it out
dropdown_demo <- function() {
  # Set our UI
  ui <- shiny::fluidPage(
    # Call our module UI
    dropdown_ui(
      id = "select",
      label = shiny::h5("Select an option:")
    ),
    # Let's create a text output to test our 
    # module
    shiny::textOutput(
      'select_text'
    )
  )
  # server
  server <- function(input, output, session) {
    # Call our server and set the choices list
    select <- dropdown_server(
      id  = "select",
      choice_list = c(
        "A",
        "B",
        "C"
      )
    )
    # Render our text output, which remember was stored in txt
    output$select_text <- shiny::renderText(select$txt)
  }
  shinyApp(ui, server)
}

dropdown_demo()
```

Alternatively, we can code it in a way that better reflects mastering shiny's documentation:

``` {r eval=F}

dropdown_ui <- function(id, label) {
  shiny::selectizeInput(
    shiny::NS(id, "select"), # changed the id to the mastering shiny method
    shiny::h5(label),
    choices = NULL, 
    options = list(allowEmptyOption = FALSE) 
  )
}

dropdown_server <- function(id, choice_list) {
  shiny::moduleServer(
    id,
    function (input, output, session) {
      shiny::updateSelectizeInput(
        session,
        id,
        choices = choice_list,
        selected = choice_list[[1]][1]
      )
      # We changed this to use a reactive container and call the select
      # input we set in the UI. It knows that we are updating this input
      # due to the id argument. 
      shiny::reactive(input$select)
    })
}


# Finally, let's test it out
dropdown_demo <- function() {
  ui <- shiny::fluidPage(
    dropdown_ui(
      id = "drop1",
      label = shiny::h5("Select an option:")
    ),
    shiny::textOutput(
      'select_text'
    )
  )
  # server
  server <- function(input, output, session) {
    letter <- dropdown_server(
      id  = "drop1",
      choice_list = c(
        "A",
        "B",
        "C"
      )
    )
    # we call the actual value using letter() to fill our text output
    # unlike the previous method
    output$select_text <- shiny::renderText(letter()) 
  }
  shinyApp(ui, server)
}

dropdown_demo()
```

::::{.notebox}
:::{.center}
**NOTE**
:::
The alternative approach improves on the code in three main ways:

- By using `NS(id, "id_name")` instead we create our internal namespace right away, whereas the original method requires tricky code at the end to link the two together `myreturn$txt <- shiny::req(eval(parse(text = paste0("input$", id))))`    
- As this module only returns one value we didn't need to return it in list format, for this reason it was simpler to call the modules return value with `letter()` as opposed to `select$txt`    
- Finally, the typical approach to calling values is by using `value()` versus the reactive call with `value`, the original doesn't follow this approach    
::::

### CSV Download Module

Next let's look at a module that allows for a CSV to be downloaded:

``` {r eval=F}
# UI set up, the typical id argument and we
# add a label argument
csv_download_ui <- function(id, label) {
  
  # Alternative method, could use the standard if we wanted
  ns <- shiny::NS(id)
  
  # for a download we just use a button
  shiny::downloadButton(
    outputId = ns('download'), # setting the id
    label = label, # our download title
    icon = shiny::icon("download") # add in a download icon
    )
  
}

# server setup, two arguments: one for the id, and one for
# the data that will be downloaded
csv_download_serv <- function(id, data_name) {
  
  # the standard copy and paste start
  shiny::moduleServer(
    id, 
    function(input, output, session){
       
      # here is our download handler
      output$download <- shiny::downloadHandler(
        # this is the format for the download handler
        filename = function() {
          # set up our file name based on the argument
          paste0(data_name, ".csv")
        },
        # this is the format for the download handler as well
        content = function(file) {
          # Write the dataset to the `file` that will be downloaded
          write.csv(eval(parse(text = data_name)), file)
        }
        
      )

  })
}

# test ------------------------------------------------------------

download_demo <- function() {
  ui <- shiny::fluidPage(
    csv_download_ui(id = "download", label = "Download csv")
  )
  # server
  server <- function(input, output, session) {
    csv_download_serv(id  = "download", data_name = "mtcars")
  }
  shinyApp(ui, server)
}

download_demo()

```

### Domain select module

This next module looks at doing something similar to the dropdown module:

``` {r eval=F}
domain_list <- list(
  "Demographics" = c("Population projection", "Housing"),
  "Factors that affect health" = c("Body mass index adult", "Body mass index youth")
)

# Setting up the UI with just an id argument
# hmm.. there 
domain_select_ui <- function(id) {
  # The other way!
  ns <- shiny::NS(id)
  # Setting this up in a container
  # this is fine but also can restrict it's resuability in the future
  shiny::fluidRow(
    # Domain selection dropdown
    shiny::selectizeInput(
      ns("domain"),
      "Select domain", # the title
      choices = names(domain_list), # initial list
      selected = names(domain_list)[1] #the initial selection
    ),
    # Second dropdown for selection
    shiny::selectizeInput(
      ns("sub_domain"), 
      "Select subdomain", # the title
      choices = NULL # empty because updating later on
    )
  )
}


## Server setup, we are looking to link our top level selection with the lower level
domain_select_server <- function(id) {
  shiny::moduleServer(
    id,
    function (input, output, session) {
      # Update HSDA selectizeInput list when
      # a new domain is selected
      shiny::observeEvent(input$domain, {
        # Look at the list and extract the sub contents
        # based on the domain
        choices <- domain_list[[input$domain]]
        # Update the sub domain list based on this selection
        shiny::updateSelectizeInput(
          session,
          "sub_domain",
          choices = shiny::isolate(choices), # I'm not sure we actually need this isolate() call.. try it without!
          server = FALSE,
          selected = choices[1]
        )
      })
      # return a list of values
      return <- shiny::reactiveValues()
      shiny::observe({
        return$domain <- shiny::req(input$domain) 
        return$sub_domain <- shiny::req(input$sub_domain)
      })
      return(return)
    })
}


## Test demo --------------------------------------------------
domain_select_demo <- function() {
  ui <- shiny::fluidPage(
    domain_select_ui(
      id = "domain_select"
    ),
    shiny::htmlOutput("domain"),
    shiny::htmlOutput("sub_domain")
  )
  # server
  server <- function(input, output, session) {
    domain <- domain_select_server(
      id  = "domain_select"
    )
    shiny::observeEvent(domain, {
      output$domain <- shiny::renderText({domain$domain})
      output$sub_domain <- shiny::renderText({domain$sub_domain})
    })
  }
  shiny::shinyApp(ui, server)
}

domain_select_demo()

```

### Sidebar select module

Lastly, we can look at one of the more complex modules that looks to create conditional panels based on the geography selection:

``` {r eval=F}

region_list <- list(
  "1 Interior" = list(
    "11 East Kootenay" = list(
      "111 Fernie" = "1110 Fernie",
      "112 Cranbrook" = "1120 Cranbrook"
    ),
    "12 Kootenay Boundary" = list(
      "121 Kootenay Lake" = "1210 Kootenay Lake",
      "122 Nelson" = "1220 Nelson"
    )
  ),
  "2 Fraser" = list(
    "21 Fraser East" = list(
      "211 Hope" = "2110 Hope",
      "212 Chilliwack" = c("2120 South Chilliwack", "2122 North Chilliwack")
    ),
    "22 Fraser North" = list(
      "221 New Westminster" = c("2211 New Westminster - East", "2212 New Westminster - Central"),
      "222 Burnaby" = c("2221 Burnaby Heights/Capital Hill", "2222 Brentwood/Willingdon/Parkcrest")
    )
  )
)

# UI set up
sidebar_select_ui <- function(id) {
  # You know the drill
  ns <- shiny::NS(id)
  # Container wrapper
  shiny::fluidRow(
    #First level of the geography selection
    shiny::selectizeInput(
      ns("geography_level"),
      "Select geography level",
      choices = list(
        "British Columbia" = "BC",
        "Health Authority" = "HA",
        "Health Service Delivery Areas" = "HSDA",
        "Local Health Area" = "LHA",
        "Community Health Service Area" = "CHSA")
    ),
    # Second level, this is for HA
    # we have a conditoinal panel as we only
    # want it to show up if we have a geography
    # level below BC
    shiny::conditionalPanel(
      condition = "input.geography_level == 'HA' ||
                   input.geography_level == 'HSDA' ||
                   input.geography_level == 'LHA' ||
                   input.geography_level == 'CHSA'",
      ns = ns,
      shiny::selectizeInput(
        ns("HA"),
        "Select HA",
        # This is hard coded, means it's not reusable, we could've
        # made this an argument to make reusable
        choices = names(region_list)
      ),
    ),
    # Third level, for HSDA
    # Anything HSDA or lower show this selection menu
    shiny::conditionalPanel(
      condition = "input.geography_level == 'HSDA'||
                   input.geography_level == 'LHA' ||
                   input.geography_level == 'CHSA'",
      ns = ns,
      shiny::selectizeInput(
        ns("HSDA"),
        "Select HSDA",
        choices = NULL, # we will update this later
        options = list(allowEmptyOption = FALSE)
      )
    ),
    # Fourth level, for LHA
    # Anything LHA or lower show this menu
    shiny::conditionalPanel(
      condition = "input.geography_level == 'LHA' ||
                   input.geography_level == 'CHSA'",
      ns = ns,
      shiny::selectizeInput(
        ns("LHA"),
        "Select LHA",
        choices = NULL,
        options = list(allowEmptyOption = FALSE)
      )
    ),
    # Fifth level, for CHSA
    # Anything CHSA how this menu
    shiny::conditionalPanel(
      condition = "input.geography_level == 'CHSA'",
      ns = ns,
      shiny::selectizeInput(
        ns("CHSA"),
        "Select CHSA",
        choices = NULL,
        options = list(allowEmptyOption = FALSE)
      )
    ),
    # A button for generating the profile
    # again this is geared for this app only, we could've just had this outside
    # to make this function a bit more generalizable
    shiny::actionButton(ns("generate_profile"), "Generate Profile")
    
  )
}
# Server set up time, standard id argument
sidebar_select_server <- function(id) {
  # The usual setup
  shiny::moduleServer(
    id,
    function (input, output, session) {
      # When we have an HA input we update the according HSDAs  
      shiny::observeEvent(input$HA, {
        shiny::updateSelectizeInput(
          session,
          "HSDA",
          choices = shiny::isolate(names(region_list[[input$HA]])), # Hmm.. again maybe we don't need this isolate(), we can test it!
          server = FALSE
        )
        # When we have an HSDA input we update the according LHAs  
        observeEvent(input$HSDA, {
          shiny::updateSelectizeInput(
            session,
            "LHA",
            # Hard coded list again.. could be an argument then we can use it for other apps!
            choices = shiny::isolate(names(region_list[[input$HA]][[input$HSDA]])),
            server = FALSE
          )
          # When we have an LHA input we update the according CHSAs  
          shiny::observeEvent(input$LHA, {
            shiny::updateSelectizeInput(
              session,
              "CHSA",
              choices = shiny::isolate(region_list[[input$HA]][[input$HSDA]][[input$LHA]]),
              server = FALSE)
          })
        })
      })
      # Create a list of values when we click the button!
      sidebar_input <- shiny::eventReactive(input$generate_profile, {
        
        list(
          geography_level = input$geography_level,
          ha = input$HA,
          hsda = input$HSDA,
          lha = input$LHA,
          chsa = input$CHSA)
      })
      # return values for use in other parts of our app!
      # Remember we actually have other ways to return lists
      # that might make this a bit more clean... interesting
      # how there are different ways to do the same thing!
      sidebar_return <- shiny::reactiveValues()
      observeEvent(input$generate_profile, {
        sidebar_return$level <- shiny::req({ sidebar_input()$geography_level })
        sidebar_return$ha <- shiny::req({ sidebar_input()$ha })
        sidebar_return$hsda <- shiny::req({ sidebar_input()$hsda })
        sidebar_return$lha <- shiny::req({ sidebar_input()$lha })
        sidebar_return$chsa <- shiny::req({ sidebar_input()$chsa })
      })
      return(sidebar_return)
    }
  )
}

# Test! ------------------------------------

sidebar_select_demo <- function() {
  
  ui <- fluidPage(
    sidebar_select_ui(id = "region_select"),
    textOutput("level"),
    textOutput("ha"),
    textOutput("hsda"),
    textOutput("lha"),
    textOutput("chsa"),
  )
  # server
  server <- function(input, output, session) {
    # If we returned a different list format than the one up there
    # we probably wouldn't need these initial `myvals <-` or `mytxt <-`. 
    #Something to consider!
    myvals <- reactiveValues()
    mytxt <- reactiveValues()
    mytxt <- sidebar_select_server(id  = "region_select")
    observe({
      myvals$level <- mytxt$level
      myvals$ha <- mytxt$ha
      myvals$hsda <- mytxt$hsda
      myvals$lha <- mytxt$lha
      myvals$chsa <- mytxt$chsa
      output$level <- renderText({ myvals$level })
      output$ha <- renderText({ myvals$ha })
      output$hsda <- renderText({ myvals$hsda })
      output$lha <- renderText({ myvals$lha })
      output$chsa <- renderText({ myvals$chsa })
    })
  }
  shinyApp(ui, server)
}

sidebar_select_demo()

```



## Additional Resources

- [Shiny Modules](https://mastering-shiny.org/scaling-modules.html#scaling-modules)
